Commutative character sheaves and geometric types for supercuspidal representations

### 2017-05-11 CJM/letter to CJM.tex
Done for this evening. Maybe.

\begin{enumerate}
\item[(1)]
\item[(2)]
\end{enumerate}

>>>>

\begin{enumerate}
\item[(1)] We are extremely grateful for this observation. We had made a mistake. Fortunately, the correction had the effect of simplifying the constructions and did not add to the length of the paper.
\item[(2)] It is quite true that it would be possible to view Section 4 as an application of our earlier paper, bypassing the the first part of the paper under review. Our main motivation for Section 1--3 was twofold: to address the `defect' in the dictionary from our earlier paper; and to extend that dictionary, as it pertains to characters, to non-commutative groups. (Although it wasn't our motivation, we were also happy to find an alternative to the stacky abelianization of Kamgarpour.) We have added a comment to this effect after Theorem 3.12, in the hopes that it will clarify this point.
\end{enumerate}

### 2017-05-11 Geometric types.tex
Done for this evening. Maybe.

We are extremely grateful to Loren Spice for explaining Yu's types for supercuspidal representations.
We also thank Masoud Kamgarpour for helpful conversations.

@@

@@

\begin{table}[ht]
\caption{Notation conversion chart.}
\begin{spacing}{1.2}
\begin{tabular}{| c|l | l | }
\hline
\text{this paper} & \text{Jiu-Kang Yu, {\it Construction of tame}} & \cite{yu:01a} \\
 & \text{{\it supercuspidal representations}} &  \\
\hline
$\oK^0$ & $\,^\circ K^0 = G^0(F)_y$ & \cite{yu:01a}*{\S 15} \\
$\oK^{i+1}$ & $\,^\circ K^{i+1} = (\,^\circ K^0) \vec{G}^{(i+1)}(F)_{y,(0, s_0, \ldots, s_{i})}$ & \cite{yu:01a}*{\S 15} \\
$\orho^0$ & $\,^\circ \rho^0$ & \cite{yu:01a}*{\S 15} \\
$\orho^{i+1}$ & $\,^\circ \rho^{i+1}$ &  \cite{yu:01a}*{\S 15} \\
$\varphi^i$ & $\phi_i\vert_{\,^\circ K^i }$ & \\
$J_{i+1}$ & $(G^i,G^{i+1})(F)_{y, (r_i, s_i)}$ &  \\
$V_{i+1}$ & $(G^i,G^{i+1})(F)_{y, (r_i, s_i)}/ (G^i,G^{i+1})(F)_{y, (r_i, s_i^+)}$ & \\ 
$V_{i+1}^\sharp$ & $(G^i,G^{i+1})(F)_{y, (r_i, s_i)}/ \ker(\widehat{\phi}_i\vert_{(G^i,G^{i+1})(F)_{y, (r_i, s_i^+)}})$ & \\
$Z_{i+1}$ & $\ker(V_{i+1}^\sharp\to V_{i+1})$ & \cite{yu:01a}*{\S 11} \\
$(f_{j+1}, j_{i+1})$ & $(f,j)$ & \cite{yu:01a}*{\S 11} \\
$\langle\ ,\ \rangle_{i+1}$ & $\langle\ ,\ \rangle$ & \cite{yu:01a}*{\S 11} \\
\hline
\end{tabular}
\end{spacing}
\label{table:notation}
\end{table}%

We can now recall how Yu uses all this to construct representations $\orho^i$ of $\oK^i$, for $i=1, \ldots, d$ and the types $(\oK^i,\orho_i)$.  \todo{Following [these sections] of Yu.}
The representations $\orho^i$ and $\orho_i$ are defined recursively.
For the base case $i=0$, set $\orho_0 \ceq \orho^0\otimes \varphi^0$; see \ref{Y1} above.
Now fix $i$.
Let $W_{i+1}$ be the Heisenberg-Weil representation of the Jacobi group $\Sp(V_{i+1})\ltimes V_{i+1}^\sharp$, whose restriction to $V_{i+1}^\sharp$ has central character $\psi_{i+1}$.
Pull-back along $h_{i+1}$ to form $h_{i+1}^*(W_{i+1})$, a representation of $\oK^i \ltimes J_{i+1}$.
Write $\inf(\orho_i)$ for the representation of $\oK^i \ltimes J_{i+1}$ obtained by pulling back $\orho^i$ along $\oK^i \ltimes J_{i+1} \to \oK^i$. 
Consider the representation
\begin{equation}\label{eq:tensor}
\orho^{i+1} \ceq h_{i+1}^*(W_{i+1}) \otimes \inf(\orho_i)
\end{equation}
of $\oK^i \ltimes J_{i+1}$.
%By construction, the character of this representation satisfies
%\begin{equation}\label{eq:tilderho}
%\trace({\tilde\rho}_{i+1})(y)= \trace(h_{i+1}^*(W_{i+1})(y)\ \trace(\orho^i)(p_i(y))\ \varphi^i(p_i(y))
%\end{equation}
%for $y \in \oK^i \ltimes J_{i+1}$.
By \cite{yu:01a}*{}, the representation $\orho^{i+1}$ of $\oK^i \ltimes J_{i+1}$ is trivial on $\oK^{i}\cap J_{i+1}$ so $\orho^{i+1}$ descends to $\oK^{i+1}$. 
%
Set $\orho_{i+1} = \orho^{i+1}\otimes \varphi^{i+1}$.
This completes the recursive definition of the Yu $(\oK^i,\orho_i)$ for $i=0, \ldots , d$.
%The character of $\orho^d$ on $\oK^d = C = J_0 \cdots J_d$ satisfies
%\[
%\trace(\rho)(k_0 \cdots k_d) = \orho^0(k_0)  \varphi^0(k_0) \cdots \varphi^d(k_d),
%\]
%for $k_i\in J_i$.

@@

\begin{proof}
For every $i=0, \ldots ,d$, set $G^i = \Gr_{R}(\underline{G}^i)$.
Recall that $G^i(\Fq) = \underline{G}^i(R) = \oK^i$, canonically.
We begin with an argument already employed in the proof of Proposition~\ref{prop:quasicharacters}.
By continuity of the quasicharacters $\varphi^i : G^i(\Fq) \to \CC^\times$, there is some $m \in \NN$ and a factorization
\[
\begin{tikzcd}
G^i(\Fq) \arrow{rr}{\varphi^i} \arrow{rd} && \EEx\\
& \underline{G}^i(R_m) \arrow{ru}[swap]{\varphi^i_m} 
\end{tikzcd}
\] 
for all $i=0, \ldots, d$.
Set $G_m^i = \Gr_m^{R}(\underline{G}^i)$.
Then $G_m^i$ is a smooth group scheme over $\Fq$ and $G_m^i(\Fq) = \underline{G}^i(R_m)$, canonically.
Using Theorem~\ref{thm:geo}, let $\cs{L}^i_m$ be a geometrization of the linear character $\varphi^i_m: G_m^i \to \CC^\times$; so
\[
\trFrob{\cs{L}^i_m} = \varphi^i_m.
\]
%Recall that the full Greenberg transform $\Gr^{R}(\underline{G}^i)$ is a group scheme over $\Fq$ such that $\Gr^{R}(\underline{G}^i)(\Fq) = \underline{G}^i(\Fq)$, canonically, and comes equipped with a morphism $\Gr^{R}(\underline{G}^i) \to \Gr^{R}_m(\underline{G}^i)$.
%Let $\cs{L}^i$ be the  Weil sheaf on $\Gr^{R}(\underline{G}^i)$ obtained from $\cs{L}^i_m$ by pullback along $\Gr^{R}(\underline{G}^i) \to \Gr^{R}_{m}(\underline{G}^i)$. 
%Then $\cs{L}^i$ is a quasicharacter sheaf, in the sense of \cite{cunningham-roe:13a}*{Def 4.3}, on $\Gr^{R}(\underline{G}^i)$ from which the quasicharacter $\varphi^i$ may be recovered.

>>>>

We are extremely grateful to Loren Spice for explaining Yu's types for supercuspidal representations.
We also thank Masoud Kamgarpour for helpful conversations, and the anonymous referee of this paper, would pointed out an error in an earlier version of the paper and helped us clarify several points.

@@

%\begin{remark}\label{rem:dictionary}
Observe that $\CCS(G)$ is a subcategory of $\CS\ab(G)$, by Definition~\ref{def:CCS}, and that Definition~\ref{def:CSab} makes it clear that there is a forgetful functor $\CS\ab(G) \to \CS(G)$, and also a forgetful functor $\CS(G)\to \Loc(G)$.
Consequently, while the category $\CCS(G)$ geometrizes characters of $G(\Fq)$ in the sense explained in Theorem~\ref{thm:geo}, it is also true that $\CS(G)$, and even $\Loc(G)$ as introduces in Section~\ref{ssec:equivariant1}, geometrizes characters of $G(\Fq)$ in the sense that for every group homomorphism $\chi : G(\Fq) \to \EEx$ there is an object $\cs{L}$ in $\Loc(G)$ such that $\trFrob{\cs{L}} = \chi$.
The point of category $\CCS(G)$ is to make this translation from characters $\chi$ to Weil local systems $\cs{L}$ as precise as possible. Theorem~\ref{thm:geo} shows that benefit of equipping Weil local systems $\cs{L}$ with the extra structure $\mu$ and $\beta$ makes the defect in this dictionary as small as it could possibly be, namely, the image $\Delta_G$ of the connecting homomorphism $G\ab(\Fq) \to \Hh^1(\Fq, G\der)$.
%\end{remark}

@@

\begin{table}[ht]
\caption{Notation conversion chart.}
\begin{spacing}{1.2}
\begin{tabular}{| c|l | l | }
\hline
\text{this paper} & \text{Jiu-Kang Yu, {\it Construction of tame}} & \cite{yu:01a} \\
 & \text{{\it supercuspidal representations}} &  \\
\hline
$\oK^0$ & $\,^\circ K^0 = G^0(F)_y$ & \cite{yu:01a}*{\S 15} \\
$\oK^{i+1}$ & $\,^\circ K^{i+1} = (\,^\circ K^0) \vec{G}^{(i+1)}(F)_{y,(0, s_0, \ldots, s_{i})}$ & \cite{yu:01a}*{\S 15} \\
$\orho^0$ & $\,^\circ \rho^0$ & \cite{yu:01a}*{\S 15} \\
$\orho^{i+1}$ & $\,^\circ \rho^{i+1}$ &  \cite{yu:01a}*{\S 15} \\
$\varphi^i$ & $\phi_i\vert_{\,^\circ K^i }$ & \cite{yu:01a}*{\S 3}\\
$J_{i+1}$ & $J^{i+1} = (G^i,G^{i+1})(F)_{y, (r_i, s_i)}$ & \cite{yu:01a}*{\S 3} \\
$V_{i+1}$ & $J^{i+1}/J^{i+1}_+ = (G^i,G^{i+1})(F)_{y, (r_i, s_i)}/ (G^i,G^{i+1})(F)_{y, (r_i, s_i^+)}$ & \cite{yu:01a}*{\S 3} \\ 
$V_{i+1}^\sharp$ & $(G^i,G^{i+1})(F)_{y, (r_i, s_i)}/ \ker(\widehat{\phi}_i\vert_{(G^i,G^{i+1})(F)_{y, (r_i, s_i^+)}})$ &  \cite{yu:01a}*{\S 4} \\
$Z_{i+1}$ & $\ker(V_{i+1}^\sharp\to V_{i+1})$ & \cite{yu:01a}*{\S 11} \\
$(f_{j+1}, j_{i+1})$ & $(f,j)$ & \cite{yu:01a}*{\S 11} \\
$\langle\ ,\ \rangle_{i+1}$ & $\langle\ ,\ \rangle$ & \cite{yu:01a}*{\S 11} \\
\hline
\end{tabular}
\end{spacing}
\label{table:notation}
\end{table}%

We can now recall how Yu uses all this to construct representations $\orho^i$ of $\oK^i$, for $i=1, \ldots, d$ and the types $(\oK^i,\orho_i)$; see \cite{yu:01a}*{\S\S 4, 15}.
The representations $\orho^i$ and $\orho_i$ are defined recursively.
For the base case $i=0$, set $\orho_0 \ceq \orho^0\otimes \varphi^0$; see \ref{Y1} above.
Now fix $i$.
Let $W_{i+1}$ be the Heisenberg-Weil representation of the Jacobi group $\Sp(V_{i+1})\ltimes V_{i+1}^\sharp$, whose restriction to $V_{i+1}^\sharp$ has central character $\psi_{i+1}$.
Pull-back along $h_{i+1}$ to form $h_{i+1}^*(W_{i+1})$, a representation of $\oK^i \ltimes J_{i+1}$.
Write $\inf(\orho_i)$ for the representation of $\oK^i \ltimes J_{i+1}$ obtained by pulling back $\orho^i$ along $\oK^i \ltimes J_{i+1} \to \oK^i$. 
Consider the representation
\begin{equation}\label{eq:tensor}
\orho^{i+1} \ceq h_{i+1}^*(W_{i+1}) \otimes \inf(\orho_i)
\end{equation}
of $\oK^i \ltimes J_{i+1}$.
%By construction, the character of this representation satisfies
%\begin{equation}\label{eq:tilderho}
%\trace({\tilde\rho}_{i+1})(y)= \trace(h_{i+1}^*(W_{i+1})(y)\ \trace(\orho^i)(p_i(y))\ \varphi^i(p_i(y))
%\end{equation}
%for $y \in \oK^i \ltimes J_{i+1}$.
By \cite{yu:01a}*{}, the representation $\orho^{i+1}$ of $\oK^i \ltimes J_{i+1}$ is trivial on $\oK^{i}\cap J_{i+1}$ so $\orho^{i+1}$ descends to $\oK^{i+1}$. 
%
Set $\orho_{i+1} = \orho^{i+1}\otimes \varphi^{i+1}$.
This completes the recursive definition of the Yu $(\oK^i,\orho_i)$ for $i=0, \ldots , d$.
%The character of $\orho^d$ on $\oK^d = C = J_0 \cdots J_d$ satisfies
%\[
%\trace(\rho)(k_0 \cdots k_d) = \orho^0(k_0)  \varphi^0(k_0) \cdots \varphi^d(k_d),
%\]
%for $k_i\in J_i$.

@@

\begin{proof}
For every $i=0, \ldots ,d$, set $G^i = \Gr_{R}(\underline{G}^i)$.
Recall that $G^i(\Fq) = \underline{G}^i(R) = \oK^i$, canonically.
We begin with an argument already employed in the proof of Proposition~\ref{prop:quasicharacters}.
By continuity of the quasicharacters $\varphi^i : G^i(\Fq) \to \CC^\times$, there is some $m \in \NN$ and a factorization
\[
\begin{tikzcd}
G^i(\Fq) \arrow{rr}{\varphi^i} \arrow{rd} && \EEx\\
& \underline{G}^i(R_m) \arrow{ru}[swap]{\varphi^i_m} 
\end{tikzcd}
\] 
for all $i=0, \ldots, d$.
Set $G_m^i = \Gr_m^{R}(\underline{G}^i)$.
Then $G_m^i$ is a smooth group scheme over $\Fq$ and $G_m^i(\Fq) = \underline{G}^i(R_m)$, canonically.
Using Theorem~\ref{thm:geo}, let $\cs{L}^i_m$ be a geometrization of the linear character $\varphi^i_m: G_m^i(k) \to \CC^\times$; so
\[
\trFrob{\cs{L}^i_m} = \varphi^i_m.
\]
%Recall that the full Greenberg transform $\Gr^{R}(\underline{G}^i)$ is a group scheme over $\Fq$ such that $\Gr^{R}(\underline{G}^i)(\Fq) = \underline{G}^i(\Fq)$, canonically, and comes equipped with a morphism $\Gr^{R}(\underline{G}^i) \to \Gr^{R}_m(\underline{G}^i)$.
%Let $\cs{L}^i$ be the  Weil sheaf on $\Gr^{R}(\underline{G}^i)$ obtained from $\cs{L}^i_m$ by pullback along $\Gr^{R}(\underline{G}^i) \to \Gr^{R}_{m}(\underline{G}^i)$. 
%Then $\cs{L}^i$ is a quasicharacter sheaf, in the sense of \cite{cunningham-roe:13a}*{Def 4.3}, on $\Gr^{R}(\underline{G}^i)$ from which the quasicharacter $\varphi^i$ may be recovered.
